|

### 2023-02-03 paper.Rmd
Fri Feb  3 13:00:09 CET 2023: update-remote.sh

```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil (2HL)", out.width="90%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
```

@@

```{r airfoil-distribution-fitness-30gens, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Airfoil (2HL)", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens/final_fit_boxplot.png")
```

@@

```{r airfoil-size-combined-30gens, echo=FALSE, fig.cap="Median Solution Size over 30 Generations - Airfoil (2HL)", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens/size_combined_median.png")
```

@@

```{r airfoil-popDiversity, echo=FALSE, fig.cap="Median Population Diversity over 30 Generations - Airfoil (2HL)", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly.png")
```
### Dynamic Adjustment of Hidden Neurons

@@

```{r airfoil-maxIndSize-params, echo=FALSE, fig.cap="Median Number of trainable Parameters over 30 Generations - Airfoil (2HL)", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/trainableParams.png")
```

@@

```{r airfoil-maxIndSize-fitness, echo=FALSE, fig.cap="Median Best Fitness over 30 Generations - Airfoil (2HL) with dynamic adjustment of regular DAE-GP", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png")
```

@@

```{r airfoil-fitness-perGen, echo=FALSE, fig.cap="Fitness over 30 Generations - Airfoil (1HL)", out.width="80%"}
knitr::include_graphics("./img/airfoil_1hl_150hn_30gens/mean_median_fitness_byGens.png")
```
... describe differences in distribution, interquartile range blablabla...

```{r airfoil-finalFitness, echo=FALSE, fig.cap="Fitness after 30 Generations - Airfoil (1HL)", out.width="80%"}
knitr::include_graphics("./img/airfoil_1hl_150hn_30gens/final_fit_boxplot.png")
```

@@

```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Alternative Pre-Strategies - Airfoil", out.width="80%"}
knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
```

@@

```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness on Test Set after 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only).png")
```

## Effect on runtime

@@

<!-- Regarding the hidden neurons, the dimension of each hidden layer, I used two different strategies: Regular DAE-GP uses the same strategy as described by earlier work (see [@dae-gp_2022_symreg] or [@dae-gp_2020_rtree]), the number of hidden neurons is dynamically set at each generation to the maximum length of all solutions inside the current training population. For pre-trained DAE-GP another strategy had to be used, I used a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population. I decided to take this approach instead of using a fixed number of hidden neurons for the fact that it simplifies the adjustment of the parameter for later experiments on different problems. It should nonetheless be mentioned that this strategy does give pre-training on average more hidden neurons per layer resulting in a more complex network (since the pre-training population is much larger than the regular population, the chances for larger individuals inside it are also higher). -->

@@

Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model. Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might be a possible way to lower the runtime of DAE-GP if further performance optimizations to the source code, especially during model sampling, can be achieved.

@@

Another disadvantage of pre-training in DAE-GP is that in order for it to show

<!-- ## Vorteile -->

<!-- Der Einsatz von Pre-Training dient: -->

<!-- * der Reduktion der Trainingsepochen pro Generation (Reduktion der Rechenleistung, insbesondere bei komplexen DAE-LSTM nützlich?) -->

<!-- ## Nachteile -->

<!-- * Erhöhung der Rechenzeit/Ressourcen: DAE-LSTMs müssen ausreichend dimensioniert sein (2+ Hidden Layer), aktuelle Publikationen nutzen nur einen einzelnen Hidden Layer^[@dae-gp_2022_symreg] ^[@daegp_explore_exploit]. -->

<!-- * Keine signifikanten Vorteile für die Qualität der gefundenen Lösungen -->

# Limitations and open Questions

## Further Questions

>>>>

```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil", out.width="90%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
```

@@

```{r airfoil-distribution-fitness-30gens, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Airfoil", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens/final_fit_boxplot.png")
```

@@

```{r airfoil-size-combined-30gens, echo=FALSE, fig.cap="Median Solution Size over 30 Generations - Airfoil", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens/size_combined_median.png")
```

@@

```{r airfoil-popDiversity, echo=FALSE, fig.cap="Median Population Diversity over 30 Generations - Airfoil", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly.png")
```
### Dynamic Adjustment of Hidden Neurons

@@

```{r airfoil-maxIndSize-params, echo=FALSE, fig.cap="Median Number of trainable Parameters over 30 Generations - Airfoil - Dynamic adjustment of regular DAE-GP", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/trainableParams.png")
```

@@

```{r airfoil-maxIndSize-fitness, echo=FALSE, fig.cap="Median Best Fitness over 30 Generations - Airfoil - Dynamic adjustment of regular DAE-GP", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png")
```

@@

```{r airfoil-fitness-perGen, echo=FALSE, fig.cap="Fitness over 30 Generations - Airfoil - Single Hidden Layer", out.width="80%"}
knitr::include_graphics("./img/airfoil_1hl_150hn_30gens/mean_median_fitness_byGens.png")
```
... describe differences in distribution, interquartile range blablabla...

```{r airfoil-finalFitness, echo=FALSE, fig.cap="Fitness after 30 Generations - Airfoil - Single Hidden Layer", out.width="80%"}
knitr::include_graphics("./img/airfoil_1hl_150hn_30gens/final_fit_boxplot.png")
```

@@

```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil - Alternative Pre-Strategies ", out.width="80%"}
knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
```

@@

```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness on Test Set after 30 Generations - Airfoil - Alternative Pre-Trategies", out.width="80%"}
knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only).png")
```

## Effect on Run-Time

@@

@@

Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model. Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might be a possible way to lower the run time of DAE-GP if further performance optimizations to the source code, especially during model sampling, can be achieved.

@@

Another disadvantage of using pre-training in DAE-GP is that, in my experiments, it required DAE-LSTM networks to use two hidden layers (see subsection [Reduced Number of Hidden Layers]). For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by [@pmlr-v9-erhan10a]. Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see [@dae-gp_2022_symreg]), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.

## Limitations and open Questions

One major limitation for the experiments conducted during the reseach phase of this thesis has been the large computational effort for experimenting with DAE-GP. Since single runs for DAE-GP have a run-time of many hours, it was not possible for me to extend the experiments to larger population sizes or more generations. Also I did not perform an extended optimization of hyperparameters (e.g by using a grid search strategy) which might yield configurations that increase the benefits of pre-training. As described in subsection [Effect on Run-Time], sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training. It could be a worthwhile endevour to explore new strategies, that
